\LARGE \bf IMU-based pose-estimation for spherical robots with limited resources

### 2021-05-23 mfi21/ieeeconf/root.tex
Andreas
new evaluation

\subsection{Orientation}
A first experiment is performed to show the overall performance of the Autogain computation for the orientation. The IMU was manually rotated in arbitrary directions, and the orientation plotted using the Madgwick filter and the Autogain filter.
As expected, the overall behavior of the presented algorithm and the Madgwick filter is very similar in the dynamic case, as shown in Figure \ref{quaternion1}.
\begin{figure}[thpb]
\includegraphics[width=\linewidth]{./graphics/AutoVsMad.eps}
\caption{Quaternion-values of the presented fusion of complementary and Madgwick filter and only the Madgwick filter}
\label{quaternion1}
\end{figure}
Therefore the more exciting part is the change between motion and standstill, as in motion, both filters use the same algorithm but differ during standstill and slow motion.
As described in Section \ref{sec:SOA}, the Madgwick filter has a rather string jitter during slow movements and standstill. 
To evaluate specifically the change between slow motion and standstill, a further experiment was performed. The results in Figure \ref{quaternionq0} show a short excerpt of the experiment with four phases of slow motion.
For external reference, we used a precise optical tracking system. 
Manually transforming the measured IMU values to the optical reference frame then
yields the above plot.
The optical system has the same shape as the presented algorithm and lacks the perturbations of the pure Madgwick filter.
To stress the same behavior without relying on absolute orientation, which will be part of further research, Figure \ref{derivative} shoes the derivative of all three approaches.
Here the change of direction due to small oscillations and the smoother fusion algorithm in comparison to the tracking is even more apparent.

\begin{figure}
\includegraphics[width=\linewidth]{./graphics/opticalOrientation.eps}
\caption{q0-quaternion-value of the presented fusion of complementary and Madgwick filter, only the Madgwick filter and an evaluation via optical tracking }
\label{quaternionq0}
\end{figure}

\begin{figure}
\includegraphics[width=\linewidth]{./graphics/derivative.eps}
\caption{Derivative of q0-quaternion-value of the presented fusion of complementary and Madgwick filter, only the Madgwick filter and an evaluation via optical tracking }
\label{derivative}
\end{figure}

@@

\begin{thebibliography}{99}
\bibitem{c2} Rossi A.P. et al.. "DAEDALUS - Descent And Exploration in Deep Autonomy of Lava Underground Structures". Wuerzburger Forschungsberichte in Robotik und Telematik (21). 978-3-945459-33-1. (2021).
\bibitem{c1} Madgwick, Sebastian O. H.. "An efficient orientation filter for inertial and inertial / magnetic sensor arrays". (2010).
\bibitem{complementary}Mahony, R. ; Hamel, T.; Pflimlin, J.-M. ."Complementary filter design on the special orthogonal group SO(3)", Proceedings of the 44th IEEE Conference on Decision and Control. (2005).
\bibitem{c3} Kok, Manon; Hol, Jeroen D.;  Schoen,Thomas B. .  "Using Inertial Sensors for Position and Orientation Estimation", Foundations and Trends in Signal Processing: Vol. 11.(2017).
\bibitem{Valencia2017ASM} Botero Valencia, J.-S. ; Garc{\'i}a, M. R. ; Ceballos, J. V. . "A simple method to estimate the trajectory of a low cost mobile robotic platform using an IMU". International Journal on Interactive Design and Manufacturing (ijidem). (2017).
\bibitem{c4}Dam, Erik B.; Koch, Martin;  Lillholm, Martin. "Quaternions, Interpolation and Animation". (1998).
\bibitem{FastInvSqrt}"Origin of Quake3's Fast InvSqrt()". beyond3d.com/content/articles/8/. Accessed: 2021-05-01.

>>>>

\subsection{Orientation}
A first experiment is performed to show the overall performance of the Autogain computation for the orientation. The IMU was manually rotated in arbitrary directions, and the orientation plotted using the Madgwick filter and the Autogain filter.
As expected, the overall behavior of the presented algorithm and the Madgwick filter is very similar in the dynamic case, as shown in Figure \ref{quaternion1}.
\begin{figure}[thpb]
\includegraphics[width=\linewidth]{./graphics/AutoVsMad.eps}
\caption{Quaternion-values of the presented fusion of complementary and Madgwick filter and only the Madgwick filter}
\label{quaternion1}
\end{figure}
Therefore the more exciting part is the change between motion and standstill, as in motion, both filters use the same algorithm but differ during standstill and slow motion.
As described in Section \ref{sec:SOA}, the Madgwick filter has a rather string jitter during slow movements and standstill. 
To evaluate specifically the change between slow motion and standstill, a further experiment was performed.
The results in Figure \ref{distance} show a short excerpt of the experiment with three phases of slow motion and rotation at second 7, 9 and 11. 
For external reference, we used a precise optical tracking system. 
The definition and proposed calculation of the distance of 2 quaternions of \cite{kuffner} was used.
The optical system matches the overall orientation of both IMU-based algorithms quite well and lacks the perturbations of the pure Madgwick filter.
The transformation of the orientation by the optical tracking system and the IMU-based algorithms yield two unknown transformations, the relative orientation of the coordinate system of IMU to optical coordinate system and how the axes are interpreted on the tracked structure itself.
Therefore, the algorithm proposed in \cite{dornaika} was used to solve these two unknown transformations on basis of the resulting data, to match the orientations of optical and imu-based data.
As this matching of the optical system happens on basis of one of the two IMU orientation data-sets, the comparison of absolute differences may be misleading, as the transformations were matched on basis of the autogain filter. 
To stress the jitter behavior of the Madgwick without relying on absolute orientation, which will be part of further research, Figure \ref{diffdistance} shows the derivative of the distance.
Here the change of direction due to small oscillations and the smoother fusion algorithm in comparison to the tracking is even more apparent.

\begin{figure}
\includegraphics[width=\linewidth]{./graphics/distance.eps}
\caption{Distance of the proposed and the Madgwick algorithm to the optical tracking system, using the distance definiton of \cite{kuffner}}
\label{distance}
\end{figure}

\begin{figure}
\includegraphics[width=\linewidth]{./graphics/diffdistance.eps}
\caption{Derivative of the distance of the proposed and the Madgwick algorithm to the optical tracking system, using the distance definiton of \cite{kuffner}}
\label{diffdistance}
\end{figure}

@@

\begin{thebibliography}{99}
\bibitem{c2} Rossi, A.P. et al., "DAEDALUS - Descent And Exploration in Deep Autonomy of Lava Underground Structures", Wuerzburger Forschungsberichte in Robotik und Telematik (21), 978-3-945459-33-1, (2021).
\bibitem{c1} Madgwick, Sebastian O. H., "An efficient orientation filter for inertial and inertial / magnetic sensor arrays", (2010).
\bibitem{complementary}Mahony, R. ; Hamel, T.; Pflimlin, J.-M.,"Complementary filter design on the special orthogonal group SO(3)", Proceedings of the 44th IEEE Conference on Decision and Control, (2005).
\bibitem{c3} Kok, Manon; Hol, Jeroen D.;  Schoen,Thomas B.,  "Using Inertial Sensors for Position and Orientation Estimation", Foundations and Trends in Signal Processing: Vol. 11.(2017).
\bibitem{Valencia2017ASM} Botero Valencia, J.-S. ; Garc{\'i}a, M. R. ; Ceballos, J. V., "A simple method to estimate the trajectory of a low cost mobile robotic platform using an IMU", International Journal on Interactive Design and Manufacturing (ijidem), (2017).
\bibitem{c4}Dam, Erik B.; Koch, Martin;  Lillholm, Martin, "Quaternions, Interpolation and Animation", (1998).
\bibitem{FastInvSqrt}"Origin of Quake3's Fast InvSqrt()", beyond3d.com/content/articles/8/, Accessed: 2021-05-01.
\bibitem {kuffner}Kuffner, James, "Effective sampling and distance metrics for 3D rigid body path planning", Proceedings - IEEE International Conference on Robotics and Automation, pp. 3993 - 3998 Vol.4, (2004).
\bibitem{dornaika}Dornaika, F.; Horaud, R., Simultaneous robot-world and hand-eye calibration, IEEE Transactions on Robotics and Automation 14(4), pp. 617-622, (1998).
